Donkeys to Monoids: A Computational Framework for Dynamic Semantics

### 2013-05-19 paper.tex
Spell check

I build a rewriting system with a well defined set rules, discussing possible semantics. I apply the framework to a large collection of sentences, showing the conversion process step by step. Finally I discuss ways to enlarge the supported fragment of English, how we can combat and interprete the inherent ambiguity we meet, and how the flexibility of Donkey Monoids relate to the classical Montague lambda structures.

@@

\begin{description}
\item[Knowledge bases] In the field of Knowledge Representation and Reasoning, they long ago discovered that formal logic provides a good language for buliding ontologies. Typical users are crime fighters or life scientists that need to draw conclusions based on complex systems\cite{malone2010modeling} where most of the known facts are only described in natural language. Today those people use mainly have to manually build large models in gui tools, but an automated approach is much wanted.

@@

\item `The donkeys are being stubborn!': This is an often used example of an `implied action' sentence and thought to have the same meaning as `Please go out and beat them!'. The best we can do here seems to be $please\_beat\_the\_doneys()$ or perhaps $\forall x(donkey(x)\rightarrow please\_beat(x))$ or even $please(\forall x(donkey(x)\rightarrow beat(x)))$.

@@

From the examples before, and perhaps from common intuition, we might find it natural to think of the sentence as two subsentences combined by `if': `if (a farmer owns a donkey) (he beats it)'. Translating the parentheses first: `A farmer owns a donkey' logifies to $\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))$, and `he beats it' translates to $beats(x,y)$. Hence the whole formula should probably be:
%
\begin{align}
\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))) \rightarrow beats(x,y)
\end{align}
%
Which is totally wrong. The $beats(x,y)$ clause ends up refering to the variables outside of their scope, and even if we try to push it inside the parentheses:
%
\begin{align}
\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y) \rightarrow beats(x,y))) \label{wrong_donkey_2}
\end{align}
%
It doesn't work either. Sentence \eqref{wrong_donkey_2}, while syntactically correct, evaluates to true as long as a single farmer beats his donkey. In fact \textbf{as long as any farmer exists} \eqref{wrong_donkey_2} evaluates to true. This is because $\exists y$ also spans over the farmer entity (and he's not a donkey) which makes the antecedent false. Clearly this is not the formula we are after.

@@

Now the above example may seem a bit contrived. Why did I have to try to combine the meaning from the obviously wrong pieces? Wasn't some other strategy more obvious? Before we go into details of such logification strategies, let's first see how easy it could have been, if we instead of first-order logic had used Visser's Donkey Monoids:
%
\begin{align}
\Bowtie \cdot \exists x \cdot farmer(x) \cdot owns(x,y) \cdot\triangle\cdot\exists y \cdot donkey(y) \cdot\triangle\cdot\Bowtie\cdot beats(x,y)
\end{align}
%
If we read $\exists$ as `a' this is basically identical to the English sentence. The donkey monoid even supports placing the transitive verb between the two referants. Clearly some magic is going on, that we will have to investigate further.

@@

We need a framework to encode our rules into, and luckily the field of computational semantics offer several solutions. Basically there are two well described ways to go: statistical vector space methods, which is very much the most popular alternative right now, and logical methods. As you might have guessed I will lean in the dirrection of the later. The most well described methods are Montague Grammars and Discourse Representation Theory (invented after the donkey sentence).

@@

We will see later how exactly the Montague Grammar obtains a full sentence formula from its reqursive process, but in any case for a computer scientist the introduction of trees is always a promising step.

The main thing to notice now is that for Montague's approach to work, it is crucial that we can always get the meaning of a composed phrase from a composition of the meaning of its subphrases. This compositional property is exactly what we saw earlier that first-order logic doesn't posses, and Montague goes through many hoops to simulate it.

@@

\item[2 POS tagger] This is the first of three Stanford NLP programs my Haskell script is going to pass the text to. Then the POS tagger tokenizes the sentences and statstically tries to figures out part of speech for each word. This is also my first source of error.

@@

\item[4 Correference Resolution] A lot has been said about how anaphora might be resolved using formal Montague Grammar like approaches. DRT is especially focused on this topic. None the less the current champion of anaphora competitions\cite{lee2011stanford}\cite{raghunathan2010multi} is the heuristical Stanford Deterministic Coreference Resolution System, the workings of which I will explain in detail.

@@

In these chapters I'm going to describe the logical theories by Visser and earlier dynamic semantists. I will introduce some addtions of my own and prove interesting properties about them. Finally, but most importantly, I will go into depth with the tools and ideas we need to write a strong logification framework.

@@

The basic idea is very simple. For each postag we define a type. E.g. $S$ has the type $Bool$ and $VP$ has the type $Entity \rightarrow Bool$, the intuition being that the VP returns true if the entity `does what the VP described' and false otherwise. An NP is given type type $VP \rightarrow Bool$ which allows it to create a scope for a variable and pass it to the VP, or whatever it sees fitting for that NP.

@@

Later in the implementation section we will look at more issues regarding to using Montague Grammars in practise. First we need to look at a very important source of ambiguity in sentences.

@@

`Weak' and `strong' are quite accepted terms in linguistics for describing different ways to interprete quantifiers. In line with information theory, we can say that weak readings are about weak information and strong readings about strong information.

@@

We have equally many reaings for the first `A', creating nine possible interpretations. This is because `a' is a very ambigious quantifier. Other quantifiers such as `every' and `all' are more clearly defined. `Some' is another example of an ambigious quantifier.

@@

The first thing we notice is that dependency graphs focus on verbs. The verbs become the heads of the sentences, and everything else becomes the ` subject', `the object' or `the something else' of the verb. In our example `he beats it' becomes an `adverbial clause modifier' of the main verb. This means it modifies the verb, here in the sence of making the main verb a condition.

Dependency graphs naturally seems to offer more semantic information than constituent trees, and hence that they would be a good intermediate step towards our goal of logification. However the first disappointment arrise as we start to look at how they are created.

@@

I've worked hard on trying to define a Montague like set of equations for Enlish, but over the dependency graph. This has however failed, mostly due to one simple reason: In a Montague grammar we expect every type of phrase to have a specific logical type. E.g. all NPs are `VP $\rightarrow$ Bool' and VPs are `Entity $\rightarrow$ Bool'. This makes it fit very nice in a programming context, where we can define a function `parseNP :: ConstituentTree $\rightarrow$ (VP $\rightarrow$ Bool)'. With the dependency approach we don't have this option. A `subj' arrow can mean many different things. Sometimes it points at a noun, sometimes at a proper noun, sometimes something else.

In conclusion, Montague like grammars for dependency graphs could be very interesting to make. It would however require some welldefined rules for what we can expect to find under each type of arrow. This was a blind road.

@@

Coreference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large iterative applications of linguistic heuristics. Of course the procedure assumes (most likely) machine learned postagging and named entity recognition, but the main algorithm is transparent and deterministic.

@@

Since this program is going to be used for research by computational lingusts, easy access to modify the code is more useful than big user interfaces and parameterization. Our interface is going to be simple text based, with an input file of text you want translated and a command you run on said file to start producing logifications.

@@

The Stanford Parser tools provide a rich Java library for interfacing directly with the system. They have classes for things like constituent trees, dependencies, correferences and more. Obviously I couldn't use any of that. Instead I ran a simple shell script they include, which outputs the relevant information in XML.

Haskell has a nice XML library in Text.XML.Light. However the constituent found by the parser were not actually in xml, but in a simple plain text format. To parse this I wrote my own parser using the great monadic Parsec library.

The next step for my code was to tag the relevant tree leafs with correct variable names, so it could go into the Montague code. As stated before the Stanford toolchain provided me with correference tagging, but unfortunately not in a great format for my needs. If a sentence said `A man and a dog, they had a shower', it would assign the same `mention' to `a man' and `a dog', but nothing to `a shower'. Luckily Haskell is quite good at such tree manipulation. One thing I made sure was that I always use distinct variables so I never have to perform $\alpha$-conversion.

@@

In terms of the semantic types we have to be careful. Types such as `Interpretation $\rightarrow$ Bool' forgets to allow for basic `errors' such as references to unbound variables. For FOL we can check this statically and not allow the syntax, but for a DPL sentence $(\exists x\cdot P(x) \cup \exists y)\cdot Q(x)$ the existance of $x$ in $Q(x)$ depends on the value of $P(x)$.

One option in case of an error is to simply return $\bot$ and say the formula evaluated to false, but perhaps we want to distinguish the cases. I have experimented with a lot of types, such as `Interpretation $\rightarrow$ Maybe Bool', which can return the value `Nothing' in case of an error. It doesn't make the formulas any nicer though. If we decide to even implement Visser's `many world' semantics as discussed earlier, we get absolutely terrible formulas that don't honour the simplicity of the logic in any way.

@@

Once I have sculpted the input text into a nice tree shape\footnote{I use the type `PosTree IWord' where:\\
\usebox{\LstBox}}, I parse it to my `montague' function. As I described in the early chapters, this applies a set of rules recursively to create a formal interpretation.

@@

The mentioned `montague' function basically passes the tree to $\llbracket\cdot\rrbracket_\text{ROOT}$ and does a bit of cleaning of the result.

@@

Another idea for logification using machine learning is to entirely skip the Montague step. The $\Bowtie$ and $\triangle$ monoid makes logic look close enough to natural language, that it starts to become tempting utilizing machine learning methods from translation. Unfortunately this idea runs into the same barrier as the above: that for superviced learning, we first need a large base of already translated sentences.

@@

Hence, as reasonable as the introduced grammar seems, we want to reduce the likelihood of conflicts as much as possible. Until now we have been very focused on the same one or two simple sentences, or whatever sentences were relevant to the features we were discussing, but now we are going to introduce a much larger set of complicated test cases. Ideally we want test cases for testing every combination of grammar rules immaginable.

In the test of this program, we used a branch of 100 sentences. Some copied from the papers referenced, others invented based on features considered or the daily life of farmers and donkeys as immagined by the author. The sentences were converted to Visser, DPL and FOL, and the meanings were checked manually to make sense.

@@

Let's start up with a very sequential example to showcase the strength of DPL. Notice how the sentences are easily concattenated with no need for scope handling.

@@

We can also handle cases where `a farmer' is introduced inside the anticedent of an `if'.

@@

A big consequence of/decision in the grammar created, is that no new entities can be introduced in the `then part' of an `if'. The reason for this is that such objects are not guarenteed to exist on the ground of the sentence. We enforce this rule using the $?_1$ function.

@@

Notice that we treat `starves' as a transitive verb, even though the sentence is clearly ambigious. In this case the phrase parser makes the decision for us.

@@

In Montague's PTQ\cite{montague1973proper} he doesn't give rules for conjunction of NPs, but only sentences and VP's. In \cite{rooth1982conjunction} and \cite{partee1983generalized} Mats Rooth and Barbara Partee discuss different issues arrising, but never reach any usable construction. Personally I've found the below to be very useful for conjunctive NPs that act as a union.

@@

A final example that really shows how DPL scope and free variables all play along together is this the case of a `non existent' antecedent. The stanford coref tagger mistakingly tags the two men in this sentence as the same, but because negation creates a closed scope, the final sentence still comes out right.

@@

A few examples of sentences that my framework already supports, but which fail due to short commings in the Stanford toolkit are `Every dog sees a cat. It chases it.' where the anaphora doesn't get resolved, and `Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo' that manages to slightly confuse the parser.

@@

It would also be interesting to return more than one result in the case of ambiguity. We have seen how weak and strong readings of the actors in the sentences can give very different results. Also conjunction of transitive verbs like `A farmer starves and beats a donkey' are inherently ambigious. Perhaps if we returned many possible logifications for each sentence, we could take the interferance to get the real meaning.

@@

At the lowest level there are a lot of trivial additions. Support in more cases for more negatives like `no' and `nobody' hasn't been explored well. Possessive pronouns is something I've struggeled with, but if we accept that `He took his jacket' translates to $\exists y\cdot owns(x,y)\cdot jacket(y)\cdot took(x,y)$ it shouldn't be hard. Also there are things like `existential there' and more kinds of SBARs like started by `then', `whom' etc. that we can add, but for the expense of additional complicated rules.

@@

Even higher up are lots of classical linguistic problems. Sentences such as `The golden mountain does not exist', `The ancient greeks thought the morning star was the evening star' have had entire papers written on them, without any definive answers. Probably we'll have to consider the use case to know how a well defined computational logic should work. Another thing at the top level is novel words and sentence structures. These will clearly require some sort of machine learning.

Finally I have mentioned that we might be able to expand the Donkey Monoids with more functionality as we sees fit during Montague rule making. The possibility for pushing control elements into the semantic of the logic is really the beauty of Visser's approach and unfortunately a road I haven't studied much in this paper. It is not hard to immagine $(Dpl,Dpl,Dpl,Dpl,Integer,Integer)$ exchanged for something like $(Dpl,...,PosTag)$ and coresponding composition rules.

\section{Acknowledgments}
I would like to thank Samson Abrahamski for introducing me to the world of Natural Language parsing and semantics and for helpful and stimulating discussions.

@@

Finnaly I wish to thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for taking donkey's welfare seriously and rescuing neglected donkeys in the UK and all over the world.

@@

\intertext{For predicates it's mostly a matter of juggling variables, values and assigments:}

@@

\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Main.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Stanford.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/ToLogic.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Dpl.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Visser.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Tests.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Utils.hs}

>>>>

I build a rewriting system with a well defined set rules, discussing possible semantics. I apply the framework to a large collection of sentences, showing the conversion process step by step. Finally I discuss ways to enlarge the supported fragment of English, how we can combat and interpret the inherent ambiguity we meet, and how the flexibility of Donkey Monoids relate to the classical Montague lambda structures.

@@

\begin{description}
\item[Knowledge bases] In the field of Knowledge Representation and Reasoning, they long ago discovered that formal logic provides a good language for building ontologies. Typical users are crime fighters or life scientists that need to draw conclusions based on complex systems\cite{malone2010modeling} where most of the known facts are only described in natural language. Today those people use mainly have to manually build large models in GUI tools, but an automated approach is much wanted.

@@

\item `The donkeys are being stubborn!': This is an often used example of an `implied action' sentence and thought to have the same meaning as `Please go out and beat them!'. The best we can do here seems to be $please\_beat\_the\_donkeys()$ or perhaps $\forall x(donkey(x)\rightarrow please\_beat(x))$ or even $please(\forall x(donkey(x)\rightarrow beat(x)))$.

@@

From the examples before, and perhaps from common intuition, we might find it natural to think of the sentence as two sub-sentences combined by `if': `if (a farmer owns a donkey) (he beats it)'. Translating the parentheses first: `A farmer owns a donkey' logifies to $\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))$, and `he beats it' translates to $beats(x,y)$. Hence the whole formula should probably be:
%
\begin{align}
\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))) \rightarrow beats(x,y)
\end{align}
%
Which is totally wrong. The $beats(x,y)$ clause ends up referring to the variables outside of their scope, and even if we try to push it inside the parentheses:
%
\begin{align}
\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y) \rightarrow beats(x,y))) \label{wrong_donkey_2}
\end{align}
%
It doesn't work either. Sentence \eqref{wrong_donkey_2}, while syntactically correct, evaluates to true as long as a single farmer beats his donkey. In fact \textbf{as long as any farmer exists} \eqref{wrong_donkey_2} evaluates to true. This is because $\exists y$ also spans over the farmer entity (and he's not a donkey) which makes the antecedent false. Clearly this is not the formula we are after.

@@

Now the above example may seem a bit contrived. Why did I have to try to combine the meaning from the obviously wrong pieces? Wasn't some other strategy more obvious? Before we go into details of such logification strategies, let's first see how easy it could have been, if we instead of first-order logic had used Visser's Donkey Monoids:
%
\begin{align}
\Bowtie \cdot \exists x \cdot farmer(x) \cdot owns(x,y) \cdot\triangle\cdot\exists y \cdot donkey(y) \cdot\triangle\cdot\Bowtie\cdot beats(x,y)
\end{align}
%
If we read $\exists$ as `a' this is basically identical to the English sentence. The donkey monoid even supports placing the transitive verb between the two referents. Clearly some magic is going on, that we will have to investigate further.

@@

We need a framework to encode our rules into, and luckily the field of computational semantics offer several solutions. Basically there are two well described ways to go: statistical vector space methods, which is very much the most popular alternative right now, and logical methods. As you might have guessed I will lean in the direction of the later. The most well described methods are Montague Grammars and Discourse Representation Theory (invented after the donkey sentence).

@@

We will see later how exactly the Montague Grammar obtains a full sentence formula from its recursive process, but in any case for a computer scientist the introduction of trees is always a promising step.

The main thing to notice now is that for Montague's approach to work, it is crucial that we can always get the meaning of a composed phrase from a composition of the meaning of its sub-phrases. This compositional property is exactly what we saw earlier that first-order logic doesn't posses, and Montague goes through many hoops to simulate it.

@@

\item[2 POS tagger] This is the first of three Stanford NLP programs my Haskell script is going to pass the text to. Then the POS tagger tokenizes the sentences and statistically tries to figures out part of speech for each word. This is also my first source of error.

@@

\item[4 Coreference Resolution] A lot has been said about how anaphora might be resolved using formal Montague Grammar like approaches. DRT is especially focused on this topic. None the less the current champion of anaphora competitions\cite{lee2011stanford}\cite{raghunathan2010multi} is the heuristical Stanford Deterministic Coreference Resolution System, the workings of which I will explain in detail.

@@

In these chapters I'm going to describe the logical theories by Visser and earlier dynamic semanticists. I will introduce some additions of my own and prove interesting properties about them. Finally, but most importantly, I will go into depth with the tools and ideas we need to write a strong logification framework.

@@

The basic idea is very simple. For each POS-tag we define a type. E.g. $S$ has the type $Bool$ and $VP$ has the type $Entity \rightarrow Bool$, the intuition being that the VP returns true if the entity `does what the VP described' and false otherwise. An NP is given type type $VP \rightarrow Bool$ which allows it to create a scope for a variable and pass it to the VP, or whatever it sees fitting for that NP.

@@

Later in the implementation section we will look at more issues regarding to using Montague Grammars in practice. First we need to look at a very important source of ambiguity in sentences.

@@

`Weak' and `strong' are quite accepted terms in linguistics for describing different ways to interpret quantifiers. In line with information theory, we can say that weak readings are about weak information and strong readings about strong information.

@@

We have equally many readings for the first `A', creating nine possible interpretations. This is because `a' is a very ambiguous quantifier. Other quantifiers such as `every' and `all' are more clearly defined. `Some' is another example of an ambiguous quantifier.

@@

The first thing we notice is that dependency graphs focus on verbs. The verbs become the heads of the sentences, and everything else becomes the ` subject', `the object' or `the something else' of the verb. In our example `he beats it' becomes an `adverbial clause modifier' of the main verb. This means it modifies the verb, here in the sense of making the main verb a condition.

Dependency graphs naturally seems to offer more semantic information than constituent trees, and hence that they would be a good intermediate step towards our goal of logification. However the first disappointment arise as we start to look at how they are created.

@@

I've worked hard on trying to define a Montague like set of equations for English, but over the dependency graph. This has however failed, mostly due to one simple reason: In a Montague grammar we expect every type of phrase to have a specific logical type. E.g. all NPs are `VP $\rightarrow$ Bool' and VPs are `Entity $\rightarrow$ Bool'. This makes it fit very nice in a programming context, where we can define a function `parseNP :: ConstituentTree $\rightarrow$ (VP $\rightarrow$ Bool)'. With the dependency approach we don't have this option. A `subj' arrow can mean many different things. Sometimes it points at a noun, sometimes at a proper noun, sometimes something else.

In conclusion, Montague like grammars for dependency graphs could be very interesting to make. It would however require some well defined rules for what we can expect to find under each type of arrow. This was a blind road.

@@

Coreference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large iterative applications of linguistic heuristics. Of course the procedure assumes (most likely) machine learned POS-tagging and named entity recognition, but the main algorithm is transparent and deterministic.

@@

Since this program is going to be used for research by computational linguists, easy access to modify the code is more useful than big user interfaces and parametrization. Our interface is going to be simple text based, with an input file of text you want translated and a command you run on said file to start producing logifications.

@@

The Stanford Parser tools provide a rich Java library for interfacing directly with the system. They have classes for things like constituent trees, dependencies, coreferences and more. Obviously I couldn't use any of that. Instead I ran a simple shell script they include, which outputs the relevant information in XML.

Haskell has a nice XML library in Text.XML.Light. However the constituent found by the parser were not actually in XML, but in a simple plain text format. To parse this I wrote my own parser using the great monadic Parsec library.

The next step for my code was to tag the relevant tree leafs with correct variable names, so it could go into the Montague code. As stated before the Stanford tool chain provided me with coreference tagging, but unfortunately not in a great format for my needs. If a sentence said `A man and a dog, they had a shower', it would assign the same `mention' to `a man' and `a dog', but nothing to `a shower'. Luckily Haskell is quite good at such tree manipulation. One thing I made sure was that I always use distinct variables so I never have to perform $\alpha$-conversion.

@@

In terms of the semantic types we have to be careful. Types such as `Interpretation $\rightarrow$ Bool' forgets to allow for basic `errors' such as references to unbound variables. For FOL we can check this statically and not allow the syntax, but for a DPL sentence $(\exists x\cdot P(x) \cup \exists y)\cdot Q(x)$ the existence of $x$ in $Q(x)$ depends on the value of $P(x)$.

One option in case of an error is to simply return $\bot$ and say the formula evaluated to false, but perhaps we want to distinguish the cases. I have experimented with a lot of types, such as `Interpretation $\rightarrow$ Maybe Bool', which can return the value `Nothing' in case of an error. It doesn't make the formulas any nicer though. If we decide to even implement Visser's `many world' semantics as discussed earlier, we get absolutely terrible formulas that don't honor the simplicity of the logic in any way.

@@

Once I have sculpted the input text into a nice tree shape\footnote{I use the type `PosTree IWord' where:\\
\usebox{\LstBox}}, I parse it to my `Montague' function. As I described in the early chapters, this applies a set of rules recursively to create a formal interpretation.

@@

The mentioned `Montague' function basically passes the tree to $\llbracket\cdot\rrbracket_\text{ROOT}$ and does a bit of cleaning of the result.

@@

Another idea for logification using machine learning is to entirely skip the Montague step. The $\Bowtie$ and $\triangle$ monoid makes logic look close enough to natural language, that it starts to become tempting utilizing machine learning methods from translation. Unfortunately this idea runs into the same barrier as the above: that for supervised learning, we first need a large base of already translated sentences.

@@

Hence, as reasonable as the introduced grammar seems, we want to reduce the likelihood of conflicts as much as possible. Until now we have been very focused on the same one or two simple sentences, or whatever sentences were relevant to the features we were discussing, but now we are going to introduce a much larger set of complicated test cases. Ideally we want test cases for testing every combination of grammar rules imaginable.

In the test of this program, we used a branch of 100 sentences. Some copied from the papers referenced, others invented based on features considered or the daily life of farmers and donkeys as imagined by the author. The sentences were converted to Visser, DPL and FOL, and the meanings were checked manually to make sense.

@@

Let's start up with a very sequential example to showcase the strength of DPL. Notice how the sentences are easily concatenated with no need for scope handling.

@@

We can also handle cases where `a farmer' is introduced inside the antecedent of an `if'.

@@

A big consequence of/decision in the grammar created, is that no new entities can be introduced in the `then part' of an `if'. The reason for this is that such objects are not guaranteed to exist on the ground of the sentence. We enforce this rule using the $?_1$ function.

@@

Notice that we treat `starves' as a transitive verb, even though the sentence is clearly ambiguous. In this case the phrase parser makes the decision for us.

@@

In Montague's PTQ\cite{montague1973proper} he doesn't give rules for conjunction of NPs, but only sentences and VP's. In \cite{rooth1982conjunction} and \cite{partee1983generalized} Mats Rooth and Barbara Partee discuss different issues arising, but never reach any usable construction. Personally I've found the below to be very useful for conjunctive NPs that act as a union.

@@

A final example that really shows how DPL scope and free variables all play along together is this the case of a `non existent' antecedent. The Stanford Coreference tagger mistakenly tags the two men in this sentence as the same, but because negation creates a closed scope, the final sentence still comes out right.

@@

A few examples of sentences that my framework already supports, but which fail due to short comings in the Stanford toolkit are `Every dog sees a cat. It chases it.' where the anaphora doesn't get resolved, and `Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo' that manages to slightly confuse the parser.

@@

It would also be interesting to return more than one result in the case of ambiguity. We have seen how weak and strong readings of the actors in the sentences can give very different results. Also conjunction of transitive verbs like `A farmer starves and beats a donkey' are inherently ambiguous. Perhaps if we returned many possible logifications for each sentence, we could take the intersection to get the real meaning.

@@

At the lowest level there are a lot of trivial additions. Support in more cases for more negatives like `no' and `nobody' hasn't been explored well. Possessive pronouns is something I've struggled with, but if we accept that `He took his jacket' translates to $\exists y\cdot owns(x,y)\cdot jacket(y)\cdot took(x,y)$ it shouldn't be hard. Also there are things like `existential there' and more kinds of SBARs like started by `then', `whom' etc. that we can add, but for the expense of additional complicated rules.

@@

Even higher up are lots of classical linguistic problems. Sentences such as `The golden mountain does not exist', `The ancient Greeks thought the morning star was the evening star' have had entire papers written on them, without any definite answers. Probably we'll have to consider the use case to know how a well defined computational logic should work. Another thing at the top level is novel words and sentence structures. These will clearly require some sort of machine learning.

Finally I have mentioned that we might be able to expand the Donkey Monoids with more functionality as we sees fit during Montague rule making. The possibility for pushing control elements into the semantic of the logic is really the beauty of Visser's approach and unfortunately a road I haven't studied much in this paper. It is not hard to imagine $(Dpl,Dpl,Dpl,Dpl,Integer,Integer)$ exchanged for something like $(Dpl,...,PosTag)$ and corresponding composition rules.

\section{Acknowledgments}
I would like to thank Samson Abramsky for introducing me to the world of Natural Language parsing and semantics and for helpful and stimulating discussions.

@@

Finally I wish to thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for taking donkey's welfare seriously and rescuing neglected donkeys in the UK and all over the world.

@@

\intertext{For predicates it's mostly a matter of juggling variables, values and assignments:}

@@

\lstinputlisting[style=bashstyle,caption=\lstname]{run.sh}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Main.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Stanford.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/ToLogic.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Dpl.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Visser.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Tests.hs}
\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Utils.hs}
